---
title: "farsdata package - vignette introduction"
author: "mmorales"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This is a simple package which provides functions helping you to explore the data given by the **National Highway Traffic Safety Administration (NHTSA)**, called **Fatality Analysis Reporting System (FARS)**. The background of the package is the final assignment of the Coursera course "Building R Packages". 

## data

Given are data of three years:  
- accident_2013.csv.bz2  
- accident_2014.csv.bz2  
- accident_2015.csv.bz2  

Each file containing the data of fatal vehicle crashes within the United States.  
An example of the head of the 2013 data file is shown below:
```{r}
# setwd("C:/Users/marco/OneDrive/061 Coursera/spec_MasteringSoftwareDevInR/03_BuildingRPackages/w2_assignment/data")

name_2013 <- make_filename(2013)
fars_2013 <- fars_read(name_2013) 
dim(fars_2013)
head(fars_2013)
```

## Workflow

To load the data, you can work with the given functions within this package:  
- `make_filename()`  
- `fars_read()`  
- `fars_summarize_years()`  
- `fars_map_state()`  

#### 1. Create filename
Depending on which year you would like to load, the filename can be accordingly created:
```{r}
name_2013 <- make_filename(2013)
name_2013
```

#### 2.1 Load data - single year
Data of one single year can be loaded via the `fars_read_years()` function and have only one year as the input. Make sure, that you have the working directory correctly
```{r}
# setwd("C:/Users/marco/OneDrive/061 Coursera/spec_MasteringSoftwareDevInR/03_BuildingRPackages/w2_assignment/data")
fars_2013 <- fars_read_years(2013)
fars_2013[[1]]
```

#### 2.2 Load data - multiple years
Same approach, but with various year as input
```{r}
# setwd("C:/Users/marco/OneDrive/061 Coursera/spec_MasteringSoftwareDevInR/03_BuildingRPackages/w2_assignment/data")
fars_3years <- fars_read_years(2013:2015)
```

#### 3. Summarizing the data
the `fars_summarize_years()` function takes the loaded data and applies a summarize function on it: in this case, it counts the fatalities by month and year:
```{r}
# setwd("C:/Users/marco/OneDrive/061 Coursera/spec_MasteringSoftwareDevInR/03_BuildingRPackages/w2_assignment/data")
fars_summary <- fars_summarize_years(2013:2015)
fars_summary
```

#### 5. Mapping fatal crashes
Finally, the `fars_map_state()` function takes the state ID and a year as an input and plots the fatality. Make sure you have the `mapdata` package loaded. 
```{r}
# setwd("C:/Users/marco/OneDrive/061 Coursera/spec_MasteringSoftwareDevInR/03_BuildingRPackages/w2_assignment/data")
require(mapdata)
fars_map_state(53, 2014)
fars_map_state(36, 2014)
```
